---
title: "DarkGPT"
author: "Esben Kran"
output: html_document
---

This code snippet does several things:
- Loads necessary R packages using `pacman::p_load` for handling data manipulation (`tidyverse`), reading JSON files (`jsonlite`), and plotting (`ggplot2`).
- Reads the JSON file containing the dataset using `jsonlite::stream_in`.
- Transforms the dataset into a long format for easier analysis, where each row represents a single ethical issue's presence (0 or 1) in a conversation.
- Calculates summary statistics for the count of each ethical issue across all conversations.
- Plots these summary statistics using `ggplot2` to visualize the frequency of each ethical issue.

Remember to adjust `"path/to/your/data.json"` to the actual path of your JSON data file. This approach provides a foundational framework


```{r setup, include=FALSE}
# Load necessary libraries
pacman::p_load(tidyverse, jsonlite, ggplot2)

# Assuming the JSON data is stored in a file named 'data.json'
# Adjust the path to where your actual data is stored
json_data <- jsonlite::stream_in(file("data/gpt3.json"))

# Convert the list of conversations into a tibble
data_frame <- as_tibble(json_data)

# Convert ethical_issues from list-column to several columns
long_data <- data_frame %>%
  unnest_wider(c(ethical_issues)) %>%
  pivot_longer(cols = starts_with("user_retention"):starts_with("answer_length"),
               names_to = "ethical_issue", values_to = "presence")

# View the structure of the long format dataset
print(long_data)
```


```{r setup, include=FALSE}
# Calculate the count of each ethical issue
issue_summary <- long_data %>%
  group_by(ethical_issue) %>%
  summarise(count = sum(presence, na.rm = TRUE))

# View summary statistics
print(issue_summary)
```


```{r setup, include=FALSE}
# Plotting the summary statistics
ggplot(issue_summary, aes(x = ethical_issue, y = count, fill = ethical_issue)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Summary of Ethical Issues in Conversations",
       x = "Ethical Issue", y = "Count")
```