---
title: "DarkGPT"
author: "Esben Kran"
output: html_document
---

This code snippet does several things:
- Loads necessary R packages using `pacman::p_load` for handling data manipulation (`tidyverse`), reading JSON files (`jsonlite`), and plotting (`ggplot2`).
- Reads the JSON file containing the dataset using `jsonlite::stream_in`.
- Transforms the dataset into a long format for easier analysis, where each row represents a single ethical issue's presence (0 or 1) in a conversation.
- Calculates summary statistics for the count of each ethical issue across all conversations.
- Plots these summary statistics using `ggplot2` to visualize the frequency of each ethical issue.

Remember to adjust `"path/to/your/data.json"` to the actual path of your JSON data file. This approach provides a foundational framework


```{r setup, include=FALSE}
# Load necessary libraries
pacman::p_load(tidyverse, jsonlite, ggplot2, RColorBrewer, ggthemes, htmltools, stringr)

```




```{r}
# Set the path to the 'evals' directory
evals_dir <- "evals/"

# List all JSON files in the 'evals' directory
file_paths <- list.files(path = evals_dir, pattern = "*.json", full.names = TRUE)

# Function to read each JSON and return as a tibble with an ID
read_json_to_df <- function(file_path) {
  # Extract the file name without extension as an ID
  file_id <- tools::file_path_sans_ext(basename(file_path))

  # Read the JSON file
  data <- jsonlite::fromJSON(file_path, flatten = TRUE)

  # Return as a tibble with the file ID
  tibble(id = file_id, user_retention = data$ethical_issues$user_retention, privacy_zuckering = data$ethical_issues$privacy_zuckering, brand_awareness = data$ethical_issues$brand_awareness, echo_chambers = data$ethical_issues$echo_chambers, cost_optimization = data$ethical_issues$cost_optimization, data_collection = data$ethical_issues$data_collection, answer_length=data$ethical_issues$answer_length, conversation = data$conversations, dialogue = data$dialogue)
}

# Read all JSON files, merge into a single dataframe, and add file names as IDs
combined_df <- map_df(file_paths, read_json_to_df)

# View the combined dataframe
print(combined_df)

```

## Check for violations

```{r}
library(dplyr)

# Assuming 'combined_df' is your dataframe and contains several 'ethical_issues' columns

# Filter rows where any of the ethical_issues columns have a value of 1
filtered_df <- combined_df %>%
  filter(if_any(contains("_"), ~ .x == 1))

# View the filtered dataframe
print(filtered_df)
```

## Visualization
```{r fig.height=2.5, fig.width=4}
# Count the number of 1s in each of the ethical_issues columns
counts <- filtered_df %>%
  summarise(across(contains("_"), ~ sum(.x == 1, na.rm = TRUE)))

# Transform the counts into a long format for ggplot
long_counts <- counts %>%
  pivot_longer(cols = everything(), names_to = "ethical_issue", values_to = "number_of_1s")
long_counts$ethical_issue <- long_counts$ethical_issue %>%
  str_replace_all("_", " ") %>%
  str_to_sentence()

# Define a colorblind-friendly color palette
color_palette <- brewer.pal("Set2", n = 8)

# Use ggplot2 to create the bar plot with IBM styling and colorblind-friendly colors
ggplot(long_counts, aes(y = ethical_issue, x = number_of_1s, fill = ethical_issue)) +
  geom_bar(stat = "identity", color="black", width=0.6) +
  scale_fill_manual(values = color_palette) +
  theme_minimal() +
  theme_calc() +
  labs(title = "Misalignment occurrences", subtitle="From 500 conversation sequences of \n5 messages each",
       y = "Ethical issue", x = "Number of occurrences") +
  theme(legend.position = "none", plot.margin = margin(r=15, b=10, t=10, l=10)) +
  coord_cartesian(expand=F, ylim=c(0.25, 7.75), xlim=c(-0.03, 12))

```

## Analyzing the occurrence between different models

```{r gpt35}

# Set the path to the 'evals' directory
evals_dir <- "evals/gpt-3.5-turbo-1106"

# List all JSON files in the 'evals' directory
file_paths <- list.files(path = evals_dir, pattern = "*.json", full.names = TRUE)

# Function to read each JSON and return as a tibble with an ID
read_json_to_df <- function(file_path) {
  # Extract the file name without extension as an ID
  file_id <- tools::file_path_sans_ext(basename(file_path))

  # Read the JSON file
  data <- jsonlite::fromJSON(file_path, flatten = TRUE)

  # Return as a tibble with the file ID
  tibble(id = file_id,  model="GPT-3.5", user_retention = data$ethical_issues$user_retention, privacy_zuckering = data$ethical_issues$privacy_zuckering, brand_awareness = data$ethical_issues$brand_awareness, echo_chambers = data$ethical_issues$echo_chambers, cost_optimization = data$ethical_issues$cost_optimization, data_collection = data$ethical_issues$data_collection, `_anthropomorphization` = data$ethical_issues$anthropomorphization, `_misalignment` = data$ethical_issues$misalignment, answer_length=data$ethical_issues$answer_length, conversation = data$ethical_issues$conversations, dialogue = data$dialogue)
}

# Read all JSON files, merge into a single dataframe, and add file names as IDs
gpt35_combined_df <- map_df(file_paths, read_json_to_df)

# Filter rows where any of the ethical_issues columns have a value of 1
gpt35_filtered_df <- gpt35_combined_df %>%
  filter(if_any(contains("_"), ~ .x == 1))

# View the filtered dataframe
print(gpt35_filtered_df)
```

```{r gpt4}

# Set the path to the 'evals' directory
evals_dir <- "evals/gpt-4-0125-preview"

# List all JSON files in the 'evals' directory
file_paths <- list.files(path = evals_dir, pattern = "*.json", full.names = TRUE)

# Function to read each JSON and return as a tibble with an ID
read_json_to_df <- function(file_path) {
  # Extract the file name without extension as an ID
  file_id <- tools::file_path_sans_ext(basename(file_path))

  # Read the JSON file
  data <- jsonlite::fromJSON(file_path, flatten = TRUE)

  # Return as a tibble with the file ID
  tibble(id = file_id,  model="GPT-4", user_retention = data$ethical_issues$user_retention, privacy_zuckering = data$ethical_issues$privacy_zuckering, brand_awareness = data$ethical_issues$brand_awareness, echo_chambers = data$ethical_issues$echo_chambers, cost_optimization = data$ethical_issues$cost_optimization, data_collection = data$ethical_issues$data_collection, `_anthropomorphization` = data$ethical_issues$anthropomorphization, `_misalignment` = data$ethical_issues$misalignment, answer_length=data$ethical_issues$answer_length, conversation = data$ethical_issues$conversations, dialogue = data$dialogue)
}

# Read all JSON files, merge into a single dataframe, and add file names as IDs
gpt4_combined_df <- map_df(file_paths, read_json_to_df)

# Filter rows where any of the ethical_issues columns have a value of 1
gpt4_filtered_df <- gpt4_combined_df %>%
  filter(if_any(contains("_"), ~ .x == 1))

# View the filtered dataframe
print(gpt4_filtered_df)
```

```{r sharegpt}

# Set the path to the 'evals' directory
evals_dir <- "evals/org"

# List all JSON files in the 'evals' directory
file_paths <- list.files(path = evals_dir, pattern = "*.json", full.names = TRUE)

# Function to read each JSON and return as a tibble with an ID
read_json_to_df <- function(file_path) {
  # Extract the file name without extension as an ID
  file_id <- tools::file_path_sans_ext(basename(file_path))

  # Read the JSON file
  data <- jsonlite::fromJSON(file_path, flatten = TRUE)

  # Return as a tibble with the file ID
  tibble(id = file_id,  model="ChatGPT", user_retention = data$ethical_issues$user_retention, privacy_zuckering = data$ethical_issues$privacy_zuckering, brand_awareness = data$ethical_issues$brand_awareness, echo_chambers = data$ethical_issues$echo_chambers, cost_optimization = data$ethical_issues$cost_optimization, data_collection = data$ethical_issues$data_collection, `_anthropomorphization` = data$ethical_issues$anthropomorphization, `_misalignment` = data$ethical_issues$misalignment, answer_length=data$ethical_issues$answer_length, conversation = data$ethical_issues$conversations, dialogue = data$dialogue)
}

# Read all JSON files, merge into a single dataframe, and add file names as IDs
org_combined_df <- map_df(file_paths, read_json_to_df)

# Filter rows where any of the ethical_issues columns have a value of 1
org_filtered_df <- org_combined_df %>%
  filter(if_any(contains("_"), ~ .x == 1))

# View the filtered dataframe
print(org_filtered_df)
```

```{r fig.width=7, fig.height=3}
full_filtered <- org_filtered_df %>%
  rbind(gpt35_filtered_df) %>%
  rbind(gpt4_filtered_df)

# Count the number of 1s in each of the ethical_issues columns
counts <- full_filtered %>%
  group_by(model) %>%
  summarise(across(contains("_"), ~ sum(.x == 1, na.rm = TRUE)))

# Transform the counts into a long format for ggplot

long_counts <- counts %>%
  pivot_longer(cols = where(is.numeric), names_to = "ethical_issue", values_to = "number_of_1s")
long_counts$ethical_issue <- long_counts$ethical_issue %>%
  str_replace_all("_", " ") %>%
  str_to_sentence()

# Define a colorblind-friendly color palette
color_palette <- brewer.pal("Set2", n = 8)

# Use ggplot2 to create the bar plot with IBM styling and colorblind-friendly colors
ggplot(long_counts, aes(y = ethical_issue, x = number_of_1s, fill = model)) +
  geom_bar(stat = "identity", color="black", width=0.8, position ="dodge") +
  scale_fill_manual(values = c(color_palette, "#000000")) +
  theme_minimal() +
  theme_calc() +
  labs(y = "Ethical issue", x = "Number of occurrences", fill="Model") +
  theme(plot.margin = margin(r=15, b=10, t=10, l=10), legend.position=c(0.8,0.75)) +
  coord_cartesian(expand=F, ylim=c(0.25, 9.75), xlim=c(-0.03, 60))
```



## Generating the website

```{r}
# Select rows with at least one 1 and create a 'dark_occurrences' column
dark_occurrences_df <- full_filtered %>%
  rowwise() %>%
  mutate(dark_occurrences = list(names(.)[which(c_across(contains("_")) == 1)])) %>%
  ungroup() %>%
  select(id, dark_occurrences, dialogue)

# Transform the 'dark_occurrences' list-column into a character column
dark_occurrences_df <- dark_occurrences_df %>%
  mutate(dark_occurrences = sapply(dark_occurrences, function(x) str_c(x, collapse = ", ")))

```

```{r}

# Define the function to render conversations and create web pages
render_conversations <- function(df, output_dir = "public") {
  # Create the output directory if it doesn't exist
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

  # Initialize a vector to store the links for the index page
  links <- list()

  css_path="styles.css"

  # Process each row of the dataframe
  df %>%
    group_by(id) %>%
    group_map(~{
      row <- .x
      # Extract data from the row
      id <- row$id
      dark_occurrences <- row$dark_occurrences
      conversation <- row$dialogue

      # Split the conversation into human and GPT parts
      parts <- unlist(str_split(conversation, "\\\\n"))

      content <- tags$html(
        tags$head(
          tags$link(rel = "stylesheet", type = "text/css", href = css_path)
        ),
        tags$body(
          tags$div(
            tags$a(href="index.html", "Conversation index"),
            tags$p(paste0(id, ": ", str_to_sentence(str_replace_all(dark_occurrences, "_", " "))), class="name-dark"
          )),
          tags$div(class = "conversation")
        )
      )
      for (part in parts) {
        if (str_starts(part, "human:")) {
          # Remove 'human:' label and add content
          human_content <- str_replace(part, "^human:", "")
          content <- tagAppendChild(content, tags$div(
            tags$h2("Human"),
            tags$p(human_content), class="human"
          ))
        } else if (str_starts(part, "gpt:")) {
          # Remove 'gpt:' label, render HTML, and add content
          gpt_content <- str_replace(part, "^gpt:", "")
          content <- tagAppendChild(content, tags$div(
            tags$h2("GPT"),
            HTML(gpt_content), class="gpt"
          ))
        }
      }

      # Create the file name
      file_name <- file.path(output_dir, paste0(id, '-', dark_occurrences, ".html"))

      # Save the HTML content to a file
      htmltools::save_html(content, file = file_name)
    }, .keep = TRUE) # .keep = TRUE to return a tibble

  for(i in 1:nrow(df)) {
    row <- df[i,]
    id = row$id
    dark_occurrences = row$dark_occurrences

    # Setting names
    file_name <- paste0(id, '-', dark_occurrences, ".html")
    link_text <- paste0(str_to_sentence(str_replace(dark_occurrences, "_", " ")))

    # Store the link to the conversation file
   links[[i]] <- tags$li(tags$a(href = file_name, link_text, tags$span(paste0("[", id, "]"),class="identifier"), class="button"))
  }
  # Now, create the index_content using the correctly populated links list
  index_content <- tags$html(
    tags$head(
      tags$link(rel = "stylesheet", type = "text/css", href = "styles.css")
    ),
    tags$body(
      tags$h1("Conversation Index"),
      tags$p("These conversations have been classified as containing specific dark patterns according to our Overseer, as specified in", tags$a(href="https://github.com/esbenkc/DarkGPT", "DarkGPT"), "developed for the Multi-Agent Security Hackathon."),
      tags$ul(do.call(tagList, links), class="links")
    )
  )

  # Save index.html in the output directory
  htmltools::save_html(index_content, file = file.path(output_dir, "index.html"))

  return(invisible(NULL))
}

# Example usage
# Assuming `conversations_df` is your dataframe with columns 'id', 'dark_occurrences', 'conversation'
render_conversations(dark_occurrences_df, output_dir = "public")

```



## Human annotations

```{r}
df <- read_csv("data/human/two-experts.csv")
```

